Concurrent Programming with CXL

### 2020-06-24 book/paper.tex
Reordered chapters

\begin{itemize}
\item Chapter~\ref{ch:cxlintro} introduces the CXL programming
language, as it provides the language for presenting synchronization
problems and solutions.
\item Chapter~\ref{ch:concurrent} illustrates the problem of
concurrent programming through a simple example in which two processes
are concurrently incrementing a counter.
\item Chapter~\ref{ch:cxlmachine} presents the
CXL virtual machine to understand the problem
underlying concurrency better.
\item Chapter~\ref{ch:critical} introduces the concept of a
\emph{critical section} and presents various flawed implementations
of critical sections to demonstrate that implementing a critical section
is not trivial.
\item Chapter~\ref{ch:peterson} introduces Peterson's Algorithm, an
elegant solution to implementating a critical section.
\item Chapter~\ref{ch:method} gives some more details on the CXL
language.
\item Chapter~\ref{ch:spinlock} introduces ``hardware'' locks
for implemented critical sections.
\item Chapter~\ref{ch:synch} presents CXL modules
that supports locks and other synchronization primitives.
\item Chapters~\ref{ch:rdwrbusy} and~\ref{ch:rdwr2} present the
abstraction and implementation of reader/writer
locks, which allow multiple processes in a critical section if they are
all only executing read-only accesses.
Chapter~\ref{ch:rdwrbusy} also talks about busy-waiting, which is
an undesirable approach to synchronize processes.
\item Chapter~\ref{ch:semaphore} introduces \emph{semaphores},
a generalization of locks
that is good not only for mutual exclusion but also for waiting for
certain application-level conditions.
\item Chapters~\ref{ch:bb} and~\ref{ch:sbs} introduce the
\emph{bounded buffer problem} (aka \emph{producer/consumer problem})
and various solutions.
\item Chapter~\ref{ch:starvation} talks about \emph{starvation}:
the problem that in some
synchronization approaches processes may not be able to get access to a
resource they need.
\item Chapter~\ref{ch:monitors} presents
\emph{monitors} and \emph{condition variables},
another approach to process synchronication.
\item Chapter~\ref{ch:deadlock} describes \emph{deadlock}
where a set of processes are indefinitely waiting for one another to
release a resource.
\item Chapter~\ref{ch:actor} presents the \emph{actor model} as an approach to
synchronization.
\item Chapter~\ref{ch:nonblocking} introduces \emph{non-blocking} or
\emph{wait-free} synchronization algorithms,
which prevent processes waiting for one another more than a bounded number of
steps.
\item Chapter~\ref{ch:barrier} describes \emph{barrier synchronization},
useful in high-performance computing applications such as parallel simulations.
\item Chapter~\ref{ch:abp} presents a problem and a solution to the distributed
systems problem of having two processes communicate reliably over an unreliable
network.
\end{itemize}

@@

>>>>

\begin{itemize}
\item Chapter~\ref{ch:cxlintro} introduces the CXL programming
language, as it provides the language for presenting synchronization
problems and solutions.
\item Chapter~\ref{ch:concurrent} illustrates the problem of
concurrent programming through a simple example in which two processes
are concurrently incrementing a counter.
\item Chapter~\ref{ch:cxlmachine} presents the
CXL virtual machine to understand the problem
underlying concurrency better.
\item Chapter~\ref{ch:critical} introduces the concept of a
\emph{critical section} and presents various flawed implementations
of critical sections to demonstrate that implementing a critical section
is not trivial.
\item Chapter~\ref{ch:peterson} introduces Peterson's Algorithm, an
elegant solution to implementating a critical section.
\item Chapter~\ref{ch:method} gives some more details on the CXL
language.
\item Chapter~\ref{ch:spinlock} introduces ``hardware'' locks
for implemented critical sections.
\item Chapter~\ref{ch:synch} presents CXL modules
that supports locks and other synchronization primitives.
\item Chapters~\ref{ch:rdwrbusy} and~\ref{ch:rdwr2} present the
abstraction and implementation of reader/writer
locks, which allow multiple processes in a critical section if they are
all only executing read-only accesses.
Chapter~\ref{ch:rdwrbusy} also talks about busy-waiting, which is
an undesirable approach to synchronize processes.
\item Chapter~\ref{ch:semaphore} introduces \emph{semaphores},
a generalization of locks
that is good not only for mutual exclusion but also for waiting for
certain application-level conditions.
\item Chapters~\ref{ch:bb} and~\ref{ch:sbs} introduce the
\emph{bounded buffer problem} (aka \emph{producer/consumer problem})
and various solutions.
\item Chapter~\ref{ch:starvation} talks about \emph{starvation}:
the problem that in some
synchronization approaches processes may not be able to get access to a
resource they need.
\item Chapter~\ref{ch:monitors} presents
\emph{monitors} and \emph{condition variables},
another approach to process synchronication.
\item Chapter~\ref{ch:deadlock} describes \emph{deadlock}
where a set of processes are indefinitely waiting for one another to
release a resource.
\item Chapter~\ref{ch:actor} presents the \emph{actor model}
and \emph{message passing} as an approach to synchronization.
\item Chapter~\ref{ch:abp} presents a problem and a solution to the distributed
systems problem of having two processes communicate reliably over an unreliable
network.
\item Chapter~\ref{ch:nonblocking} introduces \emph{non-blocking} or
\emph{wait-free} synchronization algorithms,
which prevent processes waiting for one another more than a bounded number of
steps.
\item Chapter~\ref{ch:barrier} describes \emph{barrier synchronization},
useful in high-performance computing applications such as parallel simulations.
\end{itemize}

@@

A distributed system is a concurrent system in which a collection
of processes communicate by message passing, much the same as
in the actor model.
The most important difference between distributed and concurrent
systems is that the former takes \emph{failures} into account.
In this chapter, we will consider two actors, a sender and
a receiver, that communicate by message passing over a network
that may lose messages.
